---
title: "Code for Question 4"
output: 
  html_notebook:
    number_sections: true
    toc_depth: 3
---


```{r echo = FALSE, results='hide'}
knitr::opts_chunk$set(tidy = TRUE)
getwd()
```

# Code walkthrough


## Library imports

```{r echo = TRUE, results='hide'}
library(readr)  # file input-output functions

library(stringr) # string processing functions
```

## Functions

### Defining function to get neighbours according to 4-connectivity

This function takes as input a matrix `x` and the co-ordinates of a point $(i, j)$ as `i` and `j` respectively. It then endeavours to find its four neighbours according to the four-neighbour connectivity criteria. It returns a character vector `n_collection` containing the point's neighbours' co-ordinates as comma-separated strings of form "i,j".

```{r}
find_4_n <- function(x, i, j){

    n_collection <- vector(mode = "character") # initialising empty vector to hold neighbour coordinates


  # next row same column
  n1_i <- i + 1
  n1_j <- j

  if(n1_i <= nrow(x)){ # makes sure neighbour coordinate does not come before first column
    
    if(x[n1_i, n1_j] == 1){
      n1_string <- paste(n1_i, n1_j, sep = ",") # comma-separated string of neighbour's coordinates
      n_collection <- c(n_collection, n1_string) # add the string to the neighbour collection
    }
  }

    # previous row same column
    
    n2_i <- i -1
    n2_j <- j 
    
    if(n2_i  >= 1){ 
      
      if(x[n2_i, n2_j] == 1){ # makes sure neighbour coordinate does not come before first row
        n2_string <- paste(n2_i, n2_j, sep = ",")
        n_collection <- c(n_collection, n2_string)
      }
    }

    # same row next column
    n3_i <- i
    n3_j <- j + 1 
    
    if(n3_j  <= ncol(x)){ # makes sure neighbour coordinate does not come afte last column
      
      if(x[n3_i, n3_j] == 1){
        n3_string <- paste(n3_i, n3_j, sep = ",")
        n_collection <- c(n_collection, n3_string)
      }
    }

    # same row previous column
    n4_i <- i
    n4_j <- j - 1 
    
    if(n4_j  >= 1){ # makes sure neighbour coordinate does not come before first column
      
      if(x[n4_i, n4_j] == 1){
        n4_string <- paste(n4_i, n4_j, sep = ",")
        n_collection <- c(n_collection, n4_string)
      }
    }
    return(n_collection)
}

```

### Defining function to get 4-neighbour collections

This function iterates through all the points in the matrix, and for each point gets its four-neighbour collection (or set) using `find_4_n` as a sub-routine, and accumulates the neighbour sets of each point in a list called `n_list`. 

It then looks across all the accumulated four-neighbour sets for common points within the sets, and if there are any common points it attempts to collapse these sets into a single set. The objective is to create a non-redundant collection of sets with no points in common between them, where each point is associated with only those points who are all linked with it, and they in turn are linked with others using the four-neighbour criteria. These are the four-neighbour clusters.

This function takes as an input a matrix `x` and returns the list `n_list`, where each list element is a character vector corresponding to the co-ordinates of the cluster.

The **name** of each list element is a comma-separated string of the co-ordinates of a particular point, "i,j", and the **element itself** is a character vector of points (also specified by comma-separated strings) which are linked to the point specified by the element name. All of these points correspond to a cluster.


```{r }
get_4_n_collections <- function(x){
  
  n_list <- list("blank" = NULL) # initialising list with empty element; list would hold the neighbour collections
  
  for(i in 1:nrow(x)){  # iterating over rows
    for(j in 1:ncol(x)){ # iterating over colunmns
      
      if(x[i, j] > 0){ # if coordinate value is not 0
        n_string <- paste(i, j, sep = ",")   # create comma-separated coordinate string for coordinate of point
        n_collection <- find_4_n(x, i, j)    # neighbour collection = 4 neighbours' coordinates in the form of comma-separated strings
        integration_status <- 0              # initialising default value as 0; if point present in n_list at any level, integration status should be 1; if point not present at any level, integration status should be 0
        
        for(name in names(n_list)){          # iterating over all the element names of the list (element names are comma-separated coordinates)
          if(n_string %in%  n_list[[name]]){ # if the the coordinate string is present as one of the element names
            n_list[[name]] <- union(n_list[[name]], n_collection)  # merge the neighbour collections of the current coordinate string and the element name that matches it
            integration_status <- 1                                # change integration status to 1
          } else if(length(intersect(n_collection, n_list[[name]])) > 0){
            n_list[[name]] <- union(n_list[[name]], n_collection)
            integration_status <- 1                                 # change integration status to 1
          }
        }
        
        if(integration_status == 0){                      # if integration status is still 0, 
          n_list[[n_string]] <- c(n_collection, n_string) # include the point in n_list
        }
      }
    }
  }
  
  ## cleanup; to remove unnecessary or repeated points; repeated points can occur if coordinates of one or more points occur in multiple elements in n_list, in which case those elements should rightfully be unified into a single cluster
  
  n_list <- n_list[!names(n_list) == "blank"]   # removing placeholder element
  
  n_list <- n_list[sapply(n_list, length) != 0] # removing list elements of length 0 (if any)
  
  to_remove <- vector(mode = "character")       # container for storing redundant elements
  
  for(name1 in names(n_list)){         # iterating over elements for the first time
    for(name2 in names(n_list)){       # iterating over elements again
      if(name2 != name1 & name2 %in% n_list[[name1]] & !name1 %in% to_remove){ # if name 2 occurs  components of name 1 and if it has not already been listed for removal  
         n_list[[name1]] <- union(n_list[[name1]], n_list[[name2]])  # unify components of name 2 and name 1
        to_remove <- c(to_remove, name2)  # collect name 2 for removal
      }else if(name2 != name1 & length(intersect(n_list[[name1]], n_list[[name2]])) > 0 & !name1 %in% to_remove){ # else if any of the components in name 2 occursin components of name 1 and if name 2 has not already been listed for removal 
        n_list[[name1]] <- union(n_list[[name1]], n_list[[name2]]) # unify components of name 2 and name 1
        to_remove <- c(to_remove, name2) # collect name 2 for removal
      }
    } 
  }
  
  n_list <- n_list[!names(n_list) %in% to_remove] # remove collected elements
  
  return(n_list)
}

```

### Defining function to get neighbours according to 8-connectivity

This function works in the same way as `find_4_n`, except that it attempts to find neighbours according to the eight-neighbour connectivity criteria instead of the four-neighbour connectivity criteria.

This function takes as input a matrix `x` and the co-ordinates of a point $(i, j)$ as `i` and `j` respectively. It returns a character vector containing its neighbours' co-ordinates as a single string in a comma-separated form "i,j".

```{r }

find_8_n <- function(x, i, j){
  
  n_collection <- vector(mode = "character")
  
  n1_i <- i + 1
  n1_j <- j
  
  if(n1_i <= nrow(x)){
    
    if(x[n1_i, n1_j] == 1){
      n1_string <- paste(n1_i, n1_j, sep = ",")
      n_collection <- c(n_collection, n1_string)
    }
  }
  
  n2_i <- i -1
  n2_j <- j 
  
  if(n2_i  >= 1){
    
    if(x[n2_i, n2_j] == 1){
      n2_string <- paste(n2_i, n2_j, sep = ",")
      n_collection <- c(n_collection, n2_string)
    }
  }
  
  n3_i <- i
  n3_j <- j + 1 
  
  if(n3_j  <= ncol(x)){
    
    if(x[n3_i, n3_j] == 1){
      n3_string <- paste(n3_i, n3_j, sep = ",")
      n_collection <- c(n_collection, n3_string)
    }
  }
  
  n4_i <- i
  n4_j <- j - 1 
  
  if(n4_j  >= 1){
    
    if(x[n4_i, n4_j] == 1){
      n4_string <- paste(n4_i, n4_j, sep = ",")
      n_collection <- c(n_collection, n4_string)
    }
  }
  
  n5_i <- i + 1
  n5_j <- j + 1 
  
  if(n5_i <= nrow(x) & n5_j <= ncol(x)){
    
    if(x[n5_i, n5_j] == 1){
      n5_string <- paste(n5_i, n5_j, sep = ",")
      n_collection <- c(n_collection, n5_string)
    }
  }
  
  n6_i <- i - 1
  n6_j <- j + 1 
  
  if(n6_i >= 1 & n6_j <= ncol(x)){
    
    if(x[n6_i, n6_j] == 1){
      n6_string <- paste(n6_i, n6_j, sep = ",")
      n_collection <- c(n_collection, n6_string)
    }
  }
  
  n7_i <- i + 1
  n7_j <- j - 1 
  
  if(n7_i <= nrow(x) & n7_j >= 1){
    
    if(x[n7_i, n7_j] == 1){
      n7_string <- paste(n7_i, n7_j, sep = ",")
      n_collection <- c(n_collection, n7_string)
    }
  }
  
  n8_i <- i - 1
  n8_j <- j - 1 
  
  if(n8_i >= 1 & n8_j >= 1){
    
    if(x[n8_i, n8_j] == 1){
      n8_string <- paste(n8_i, n8_j, sep = ",")
      n_collection <- c(n_collection, n8_string)
    }
  }
  
  return(n_collection)
}

```


### Defining function to get 8-neighbour collections

This function works in the same way as `get_4_n_collections`; however, it uses `find_8_n` as a subroutine instead of `find_4_n`. It takes as an input a matrix `x` and outputs a list `n_list`. 


```{r }
get_8_n_collections <- function(x){
  n_list <- list("blank" = NULL)
  
  for(i in 1:nrow(x)){
    for(j in 1:ncol(x)){
      if(x[i, j] > 0){
        
        n_string <- paste(i, j, sep = ",")
        n_collection <- find_8_n(x, i, j)
        integration_status <- 0
        
        for(name in names(n_list)){
          if(n_string %in%  n_list[[name]]){
            integration_status <- 1
            n_list[[name]] <- union(n_list[[name]], n_collection)
          } else if(length(intersect(n_collection, n_list[[name]])) > 0){
            integration_status <- 1
            n_list[[name]] <- union(n_list[[name]], n_collection)
          }
        }
        
        if(integration_status == 0){
          n_list[[n_string]] <- c(n_collection, n_string)
        }
      }
    }
  }
  
  n_list <- n_list[!names(n_list) == "blank"]
  
  n_list <- n_list[sapply(n_list, length) != 0]
  
  to_remove <- vector(mode = "character")

  for(name1 in names(n_list)){
    for(name2 in names(n_list)){
      if(name2 != name1 & name2 %in% n_list[[name1]] & !name1 %in% to_remove){
        n_list[[name1]] <- union(n_list[[name1]], n_list[[name2]])
        to_remove <- c(to_remove, name2)
      }else if(name2 != name1 & length(intersect(n_list[[name1]], n_list[[name2]])) > 0 & !name1 %in% to_remove){
        n_list[[name1]] <- union(n_list[[name1]], n_list[[name2]])
        to_remove <- c(to_remove, name2)
      }
    } 
  }
  
  n_list <- n_list[!names(n_list) %in% to_remove]
  
  return(n_list)
}
```


### Defining function to label the clusters

Once a non-redundant collection of interlinked neighbours (i.e., a clusters) has been generated, this function labels the points in each cluster with a unique integer label. It takes as an input the matrix to be labelled, `x`, and the a list `n_list`, each element of which contains the points corresponding to the a cluster in the matrix according to either the 4-connectivity or 8-connectivity criteria. It returns the labelled matrix `lab_mat`, with the clusters labelled by unique integer labels.

Note that this function uses within itself the function `str_split` from the `stringr` package, to split the comma-separated co-ordinates of the cluster points along the comma. [Link to the function documentation.](https://stringr.tidyverse.org/reference/str_split.html)

```{r }

label_clusters <- function(x, n_list){
  
  lab_mat <- x                  # create matrix to be labelled as a copy of the unlabelled matrix
  
  for(i in seq_along(n_list)){  # iterate through n_list indices
    cluster_label <- i          # use index as integer label for current cluster
    n_collection <- n_list[[i]] # get indices of points in cluster as comma-separated strings from within n_list
    n_collection_coord <- str_split(n_collection, pattern = ",") # split the string along the comma to get a list with each element being a character vector of two strings, the strings being the row and column indices of the points
    
    for(i in seq_along(n_collection_coord)){  # iterate along the indices of the list
      current_coords <- as.integer(n_collection_coord[[i]]) # covert the coordinate strings into integers
      lab_mat[current_coords[1], current_coords[2]] <- cluster_label # label the matrix element using the coordinates with the previously-defined cluster label
    }
  }
  return(lab_mat)
}

```



## Example walk-through

Consider the following matrix y, which was given in the statement of the original problem. 

```{r }

y <- rbind(c(0, 0, 1, 1),
           c(1, 1, 0, 0),
           c(0, 0, 0, 0),
           c(0, 1, 1, 0),
           c(0, 1, 1, 1)
           )
print(y)
```

Getting the four neighbours of a particular point in `y`. Let us pick point (4, 3) and look at its four neighbours.

```{r }
find_4_n(x = y, i = 4, j = 3)
```

Getting the eight neighbours of a particular point in `y`. Let us pick point (4, 3) and look at its eight neighbours.

```{r }
find_8_n(x = y, i = 4, j = 3)
```

Getting four-neighbour collections.

```{r }
n4_list <- get_4_n_collections(y)
print(n4_list)
```

Getting eight-neighbour collections.

```{r }
n8_list <- get_8_n_collections(y)
print(n8_list)
```

Four-neighbour labelling.

```{r }
lab4_mat <- label_clusters(y, n4_list)

lab4_mat
```

Eight-neighbour labelling.

```{r }
lab8_mat <- label_clusters(y, n8_list)

lab8_mat
```



# Actual solution to Question 4

## Read input file

```{r echo = FALSE, results = "hide"}
curDir <- getwd()
list.files(curDir, include.dirs = FALSE)
```

```{r}
curDir <- getwd()
list.dirs(curDir, recursive = FALSE, full.names = FALSE)
```


```{r}
list.files(path = file.path(curDir, "Question 4")) 
```

```{r}
mat <- read_tsv(file.path(curDir, "./Question\ 4/input_question_4"), col_names = F)

mat <- as.matrix(mat)

mat
```

### Four neighbour labelling


```{r }
n4_list <- get_4_n_collections(mat)
n4_list
```

```{r}
lab4_mat <- label_clusters(mat, n4_list)

lab4_mat
```


### Eight-neighbour labelling

```{r }
n8_list <- get_8_n_collections(mat)
n8_list
```


```{r }
lab8_mat <- label_clusters(mat, n8_list)
lab8_mat
```

### Writing outputs to files

Since the question specifies that either of the four- or eight-connectivity labelled matrices is a valid answer, the four connectivity matrix will be written to the required output file. However, the eight connectivity matrix will be provided in a separate output file in case it needs to be checked.

Writing four-connectivity output.

```{r }
write_delim(x = as.data.frame(lab4_mat), path = file.path(curDir, "Question 4", "output_question_4"),  delim = " ", col_names = FALSE)
```

Writing eight-connectivity output.

```{r }
write_delim(x = as.data.frame(lab8_mat), path = file.path(curDir, "Question 4", "output_question_4_eight_connectivity.txt"),  delim = " ", col_names = FALSE)
```

